---
title: "Covid91 vaccine study Final2024"
author: "Citlali Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Males

$H_0$: The covid vaccine does not have an impact on males     
$H_A$: The covid vaccine does have an impact on males 

## Methods

Infected is a categorical variable     
Treatment is categorical variable     
The helper file I used is Cat~Cat 

## Descriptive Results

### Graphical Descriptive Results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=males)
barchartGC(~infected +treatment, data=males,type="percent")

```
     
This graph indicates that males who received the vaccine are less likely to get covid-19 compared to those who got a placebo.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~treatment +infected, data=males)
rowPerc(table1)
colPerc(table1)
```
     
This numerical summary shows that 74% of males who received the placebo got covid-19, whereas only 26% of those who received the vaccine did.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds ratio of 2.81:1 which means that the individuals who took the placebo are 2.81 times more likely to contract covid-19 compared to those who got the vaccine.          

The p-value for the chi-square test is 0.0000000000000002, the fishers exact test yields a p-value of 0.000000000000002. Therefore, we reject the null hypothesis. 

### Conclusion

The graph shows that the men who were vaccinated had a significantly lower risk of getting covid19 compared to those that received a placebo. The data shows that 74% of males who received the placebo contracted covid19, while only 26% of those who were vaccinated did. The odds ratio of 2.81:1 indicates that those who received the placebo are 2.81 times more likely to contract COVID-19 than those who were vaccinated. The chi-square test yielded a p-value of 0.00000000000000002, and Fisher's exact test produced a p-value of 0.00000000000000002, both of which are well below 0.05, showing significant results.       

### Females

$H_0$: The covid vaccine does not have an impact on females      
$H_A$: The covid vaccine does have an impact on females

## Methods

Sex is a categorical variable     
Treatment is categorical variable     
The helper file I used is Cat~Cat 

## Descriptive Results

### Graphical Descriptive Results

```{r}
femaleData = filter (FinalData,sex == "F")
```

```{r}
females=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=females)
barchartGC(~infected +treatment, data=females,type="percent")

```

This graph indicates that females who received the vaccine are less likely to contract covid19 compared to females who received the placebo.

### Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

This numerical summary shows that 65% of females who received the placebo were infected with covid-19, whereas 35% of females who received the vaccine got the virus.

## LGBTQ



## Druggies


# Overall Results and Conclusions