---
title: "Covid91 vaccine study Final2024"
author: "Citlali Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Males

$H_0$: The covid vaccine does not have an impact on males     
$H_A$: The covid vaccine does have an impact on males 

## Methods

Infected is a categorical variable     
Treatment is categorical variable     
The helper file I used is Cat~Cat 

## Descriptive Results

### Graphical Results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=males)
barchartGC(~infected +treatment, data=males,type="percent")

```
     
This graph indicates that males who received the vaccine are less likely to get covid19 compared to those who got a placebo.

### Numerical Descriptive

```{r}
table1 <- xtabs(~treatment +infected, data=males)
rowPerc(table1)
colPerc(table1)
```
     
This numerical summary shows that 74% of males who received the placebo got covid19, whereas only 26% of those who received the vaccine did.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds ratio of 2.81:1 which means that the individuals who took the placebo are 2.81 times more likely to contract covid19 compared to those who got the vaccine.          

The p-value for the chi-square test is 0.000000000000002, the fishers exact test yields a p-value of 0.0000000000000002. Therefore, we reject the null hypothesis. 



### Females

$H_0$: The covid vaccine does not have an impact on females      
$H_A$: The covid vaccine does have an impact on females

## LGBTQ



## Druggies


# Overall Results and Conclusions